---
title: "Simulation Modelling Techniques - Course Work - 2nd Year|2nd Semester."
author: "Ravindu Wellalage"
date: "2024-04-01"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# import necessary libraries

## includes a collection of packages for data manipulation and visualization.
```{r}
library(tidyverse)
```

## discrete-event simulation package.
```{r}
library(simmer) 
```

## for visualizing simulation results.
```{r}
library(simmer.plot) 
```

## for arranging multiple plots on one page.
```{r}
library(gridExtra) 
```

# Question - (1)
```{r}
set.seed(2024) 

env<- simmer("New call centre") 

customer <- trajectory("Customer") %>%
  seize("CSR-Employee", 1) %>%
  timeout(function() rexp(n=1, rate=1/3))  %>%
  release("CSR-Employee", 1)
 
env%>% 
  add_resource("CSR-Employee", 6) %>%
  add_generator("Customer", customer,function() rpois(n=1 , lambda = 1))

env %>%
  run(until=1440)
```

## Plotting relevant graphs for employee usage and waiting time representation
```{r}

p1 <- plot(get_mon_resources(env),
           metric = "usage",
           items ="server",step = T)

p2 <- plot(get_mon_arrivals(env), metric = "waiting_time")

grid.arrange(p1,p2,ncol=1)
```

## Plotting Resourcw usage and evalution time graphs
```{r}
plot(get_mon_resources(env), matric="utilization")
plot(get_mon_arrivals(env), matric="utilization")
plot(get_mon_resources(env), matric="usage", item="server")
```
# Question - (2)

## Part A 

### Generating 7000 Random Values for power dissipated in a resistor
```{r}
# Set the reproducibility
set.seed(123)
n <- 7000

voltage <- rnorm(n, mean = 12, sd= 2)
resistence <- rnorm(n, mean = 8, sd =1)

# calculate power dissipated (p)

power <- voltage^2 / resistence
```
## Part B

### Creating scatter plot to visualize the relationship between power and voltage

```{r}
plot(voltage, power, xlab="Voltage(V)", ylab="Power(W)",
     main = "Relationship between power and voltage", col = "green", pch = 16)

# Add a regression line to plot

abline (lm(power ~ voltage),col="red")

```
### Note: As voltage increases, power tends to increase as well

## Part C

### Finding the approximate mean and variance of power (p).
```{r}
power_mean <- mean(power)
power_varience <- var(power)

cat("Approximate power mean: ",power_mean, "W\n")
cat("Approximate power varience: ",power_varience,"W^2\n")


```
## Part D

### Estimate the probability that the power (p) is greater than 20W.

```{r}
prob <- mean(power>20)
cat("Probability that power is grater than 20W: ", prob)
```
# Question - (3)

## Load ggplot2 to plotting graphs
```{r}
library(ggplot2)
```


```{r}
  samples <- matrix(runif(500 * 50, min=-5, max=5), nrow=500)
  
  # b) Calculate the sample means
  sample_means <- rowMeans(samples)
  
  # c) Plot a histogram of the sample means
  hist(sample_means, breaks=30, main=paste("Histogram of Sample Means (n=", 50, ")"),
       xlab="Sample Mean", ylab="Frequency", xlim=c(-5,5))
```

## Define the simulation function Part (a,b,c)
```{r}
simulate_sample_means <- function(N, n) {
  # a) Generate N samples of size n from a Uniform[-5,5] distribution
  samples <- matrix(runif(N * n, min=-5, max=5), nrow=N)
  
  # b) Calculate the sample means
  sample_means <- rowMeans(samples)
  
  # c) Plot a histogram of the sample means
  hist(sample_means, breaks=30, main=paste("Histogram of Sample Means (n=", n, ")"),
       xlab="Sample Mean", ylab="Frequency", xlim=c(-5,5))

  # Return the sample means vector
  return(sample_means)
}

```

## e) Run the function for n=10,15,30,50 and plot histograms

### Set up plotting area for 4 plots
```{r}
par(mfrow=c(2,2)) 
for (sample_size in c(10, 15, 30, 50)) {
  simulate_sample_means(N=500, n=sample_size)
}
```


# Question - (4)

## Part - A 

### assuming you want 100 steps in the walk
```{r}

set.seed(39)
 n <- 100 
 numRealizations <- 5 # Number of realizations
 random_walks <- matrix(nrow = numRealizations, ncol = n)
```

### Simulate the random walk in 5 realizations

```{r}
 for (i in 1:numRealizations) {
 X <- sample(c(-1, 1), n, replace = TRUE, prob = c(0.5, 0.5))
 S <- cumsum(X)
 plot(S, type = "l", main = paste("Realization",i))
 random_walks[i, ] <- c(0, head(S, -1))

 }
```
## Simulating Random walks in one graph
###Plot all the realizations on one graph
```{r}
matplot(t(random_walks), type = 'l', lty = 1, 
        xlab = "Step count", ylab = "Random Walk Value positions", 
        main = "Five Realizations of the Random Walk")
legend("topright", legend = paste("Realization", 1:numRealizations), 
       col = 1:numRealizations, lty = 1)
```



